---
title: "*Danaus plexippus* haplotype and Danainae phylogenetics"
author: "Chris Hamm"
date: "`r format(Sys.Date())`"
output:  
      html_document:
        keep_md: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

Set the seed and load the required packages, then save the version of the packages
```{r preliminaries}
set.seed(89623578)
library("phytools")
library("geiger")
library("phytools")

(sess <- sessionInfo())
```

Load the phylogeny created by MrBayes:
```{r load_phylo}
Dan <- read.nexus("Data/Bayes/Danaini_cat.nex.con.tre")
plot(Dan) # Take a look at the tree
is.rooted(Dan) # This is unrooted, we'll need to correct that.
is.ultrametric(Dan) # Will need to make ultrametric as well
```
Please note, the code to run MrBayes (and the output) is located in the GitHub repo along with the code and output from `chromEvol`.


We'll use the penalized method to make the tree ultrametric
```{r make_ultra}
lam <- 10^(-1:6)
cv <- sapply(lam, function(x) sum(attr(chronopl(Dan , lambda = x, CV = TRUE), "D2")))
cv
plot(x = lam, y = cv, pch = 19, ylab = "cross-validation score", xlab = expression(paste(lambda)), las = 1, cex = 1.5) # lowest CV is alpha 1e6, suggests that branches each have different rate
```

```{r ultra_root}
Dan2 <- chronopl(phy = Dan, lambda = 1e6, CV = TRUE, eval.max = 1e3, iter.max = 1e4)
is.ultrametric(Dan2)
Dan2 <- root(phy = Dan2, outgroup = 44, resolve.root = TRUE)
is.rooted(Dan2)
```

Read in the covariate data (the karyotypes from the literature)
```{r read_data}
Dan_data <- read.csv("Danainae_karyotypes2.csv", header = TRUE, row.names = 1)
str(Dan_data)
Dan_data
```

Now we remove the taxa in the tree but not in the data, and vice versa.
```{r prune}
Dan_norm <- treedata(phy = Dan2, data = Dan_data, sort = TRUE, warnings = TRUE)

# Now need to sort the data to make sure they are in the same order as the tip labels.
Covs <- Dan_norm$data[Dan_norm$phy$tip.label,, drop = FALSE]
```

Here is figure 1 from the pre-print
```{r fig1}
Covs2 <- as.matrix(Covs)[, 1]
seq1 <- seq(from = 20, to = 36)
ancs <- c(31, 31, 31, 29, 31, 31, 31, 30, 30, 30, 30, 28, 33, 30, 30, 31, 31)
sa <- cbind(seq1, ancs)

plot.phylo(Dan_norm$phy, align.tip.label = TRUE, type = "phylogram", use.edge.length = TRUE, edge.width = 2, root.edge = TRUE, no.margin = TRUE, adj = 0)
# tiplabels(text = Covs2, adj = -9.3, frame = "none", bg = "white")
tiplabels(text = Covs2, adj = 1.2, frame = "rect", bg = "white")
nodelabels(text = sa[, 2], frame = "rect", bg = "white")
```